Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-02-07 paper/zhang-cook-laa-langrene-menendez.Rmd
fix captions and regenerate plots

```{r tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the steps in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation."}
include_graphics(here::here("img/tour-path.png"))
```

@@

```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \\code{boa6}, using the holes index. Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8."}
```

@@

```{r toy-interp, fig.cap = "Examing the index values as the optimisation progress for two optimisers using trace plot. The lines indicate the interpolation and the dots indicate a new target is generated."}
```

@@

```{r toy-pca, fig.cap = "The PCA plot of 1D projection problem on the 5-variable dataset, \\code{boa5}, using the holes index with two optimisers: SA and PD. All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases."}
```

@@

```{r toy-pca-animated, fig.cap = "Six selected frames from the animated PCA plots in the previous figure. With animation, the start and end position of each optimisation is easier to identify.  A full video of this aimation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
```

@@

```{r toy-tour, fig.cap="Six selected frames from the tour animation for viewing the same two optimisations as previous two figures. The tour animation allows for an appreciation of the search bases in the original high dimensional space. A full video of this aimation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}"}
```

@@

```{r interruption, fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index, optimised by optimiser SA."}
```

@@

```{r polish, fig.height= 5, fig.cap = "Trace plot of 2D projection on \\code{boa6} data with holes index, optimised first by SA and then by polish search. The polish step results in clearer cut on the edges of the four clusters in the data structure."}
```

@@

```{r noisy-better-geo, fig.height = 4, fig.cap = "Trace and PCA plot of 1D projection on 5\\-variable dataset, \\code{boa5}, with \\code{norm\\_kol} index, optimised by all three optimisers. Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretial best."}
```

@@

```{r kol-result, fig.height = 4, fig.cap="PCA plot of 1D projection on the 6\\-variable data, \\code{boa6}, with normal kolmogorov index. Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates. The local optimum (x), is presented in this experiment, along with the global optimum (*). The final basis is larger and more opaque than the start basis for each path and color represents whether the global or local optimum is found, after polishing."}
```

@@

```{r flip-sign, fig.cap = "Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463. On the left panel, one target basis is generated on the other half of the basis space and the interpolation is unsatisfactory. After implementing an orientation check on the determinant, this phenomenon disappears on the right panel."}
```

>>>>

```{r tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the steps in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the optimisation algorithms, whilst the white frames are constructed between two blue frames by geodesic interpolation."}
include_graphics(here::here("img/tour-path.png"))
```

@@

```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \\code{boa6}, using the holes index. Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis."}
```

@@

```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right) using trace plot.The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. The lines indicate the interpolation and the dots indicate new targets generated by the optimisers. The interpolation with both optimisers is smooth while the on the panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130."}
```

@@

```{r toy-pca, fig.cap = "The PCA-projected basis space with two search paths optimised by SA and PD. The holes index is optimised for the 1D projection problem on the five-variable dataset, \\code{boa5}. All the bases in PD has been flipped for easier comparison of the final bases found by the two optimisers and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
```

@@

```{r toy-pca-animated, fig.cap = "Six frames selected from the animation of search paths in the PCA-projected basis space. With animation, the start and end position of each optimisation routine is clearly identified. The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504242845/b73f37175a}"}
```

@@

```{r toy-tour, fig.cap="Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}. The full video of the animation can be found at \\url{https://vimeo.com/user132007430/review/504328122/9be84db563}"}
```

@@

```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data with the holes index, optimised by SA. On the left panel, basis with higher index value is found during the interpolation but not used. The interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel."}
```

@@

```{r polish, fig.height= 5, fig.cap = "The interpolation of a 2D projection problem on \\code{boa6} data with the holes index being first optimised by SA and then then polished. The projection by the final baiss of each algorithm is also presented. Four clusters in the data are detected by SA while the polish step produces a clearer cut on the edge.."}
```

@@

```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \\code{boa5}. Optimiser PD fails to optimise this non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretical best."}
```

@@

```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space with two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7. The optimisation is done on the six-variable data, \\code{boa6} with 1D projecton index $I^{nk}(n)$. A local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*). The final basis is larger and more opaque than the start basis in each path and color represents whether the global or local optimum is found, after polishing."}
```

@@

```{r flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis. Optimisation is conducted on the 1D projection index, $I^{nk}(n)$, in boa6 data using SA with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated on the right half of the basis space and the interpolator crosses the origin to perform the interpolation, wbhich is undesirable. An orientation check is implemented to correct the sign of the target basis if the it has an opposite oreintation to the current basis. The right panel shows the same interpolation after implementing the orientation check and the undesirable interpolation disappears."}
```

### 2021-02-07 paper/zhang-cook-laa-langrene-menendez.tex
fix captions and regenerate plots

\caption[An illustration for demonstrating the steps in a tour path]{An illustration for demonstrating the steps in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are found by an optimisation algorithm iteratively whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
\end{figure}
\end{Schunk}

@@

\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection of a six-variable dataset, \code{boa6}, using the holes index. Both optimisers find the final basis with similar index value while it takes SAJO more iterations because the algorithm allows a probabilistic acceptance of bases with lower index value, as observed in iteration 4 and 6-8.}\label{fig:toy-search}
\end{figure}
\end{Schunk}

@@

\caption[Examing the index values as the optimisation progress for two optimisers using trace plot]{Examing the index values as the optimisation progress for two optimisers using trace plot. The lines indicate the interpolation and the dots indicate a new target is generated.}\label{fig:toy-interp}
\end{figure}
\end{Schunk}

@@

\caption[The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers]{The PCA plot of 1D projection problem on the 5-variable dataset, \code{boa5}, using the holes index with two optimisers: SA and PD. All the bases in PD has been flipped and a grey dashed line has been annotated to indicate the symmetry of two starting bases.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

@@

\caption{Six selected frames from the animated PCA plots in the previous figure. With animation, the start and end position of each optimisation is easier to identify.  A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
\end{figure}
\end{Schunk}

@@

\caption{Six selected frames from the tour animation for viewing the same two optimisations as previous two figures. The tour animation allows for an appreciation of the search bases in the original high dimensional space. A full video of this aimation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}}\label{fig:toy-tour}
\end{figure}
\end{Schunk}

@@

\caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised by optimiser SA.}\label{fig:interruption}
\end{figure}
\end{Schunk}

@@

\caption[Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search]{Trace plot of 2D projection on \code{boa6} data with holes index, optimised first by SA and then by polish search. The polish step results in clearer cut on the edges of the four clusters in the data structure.}\label{fig:polish}
\end{figure}
\end{Schunk}

@@

\caption[Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimisers]{Trace and PCA plot of 1D projection on 5\-variable dataset, \code{boa5}, with \code{norm\_kol} index, optimised by all three optimisers. Optimiser PD fails to optimise the non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretial best.}\label{fig:noisy-better-geo}
\end{figure}
\end{Schunk}

@@

\caption[PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index]{PCA plot of 1D projection on the 6\-variable data, \code{boa6}, with normal kolmogorov index. Two optimisers: SA and SAJO and two search sizes: 0.5 and 0.7 are used in the simulation for 20 replicates. The local optimum (x), is presented in this experiment, along with the global optimum (*). The final basis is larger and more opaque than the start basis for each path and color represents whether the global or local optimum is found, after polishing.}\label{fig:kol-result}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463]{Comparison of the basis space before and after reconciling the orientation of the current and target basis for 1D projection of boa6 data, optimised by SA with seed 2463. On the left panel, one target basis is generated on the other half of the basis space and the interpolation is unsatisfactory. After implementing an orientation check on the determinant, this phenomenon disappears on the right panel.}\label{fig:flip-sign}
\end{figure}
\end{Schunk}

>>>>

\caption[An illustration for demonstrating the steps in a tour path]{An illustration for demonstrating the steps in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the optimisation algorithms, whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
\end{figure}
\end{Schunk}

@@

\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \code{boa6}, using the holes index. Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis.}\label{fig:toy-search}
\end{figure}
\end{Schunk}

@@

\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right) using trace plot.The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}. The lines indicate the interpolation and the dots indicate new targets generated by the optimisers. The interpolation with both optimisers is smooth while the on the panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130.}\label{fig:toy-interp}
\end{figure}
\end{Schunk}

@@

\caption[The PCA-projected basis space with two search paths optimised by SA and PD]{The PCA-projected basis space with two search paths optimised by SA and PD. The holes index is optimised for the 1D projection problem on the five-variable dataset, \code{boa5}. All the bases in PD has been flipped for easier comparison of the final bases found by the two optimisers and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

@@

\caption{Six frames selected from the animation of search paths in the PCA-projected basis space. With animation, the start and end position of each optimisation routine is clearly identified. The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504242845/b73f37175a}}\label{fig:toy-pca-animated}
\end{figure}
\end{Schunk}

@@

\caption{Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}. The full video of the animation can be found at \url{https://vimeo.com/user132007430/review/504328122/9be84db563}}\label{fig:toy-tour}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data with the holes index, optimised by SA]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data with the holes index, optimised by SA. On the left panel, basis with higher index value is found during the interpolation but not used. The interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel.}\label{fig:interruption}
\end{figure}
\end{Schunk}

@@

\caption[The interpolation of a 2D projection problem on \code{boa6} data with the holes index being first optimised by SA and then then polished]{The interpolation of a 2D projection problem on \code{boa6} data with the holes index being first optimised by SA and then then polished. The projection by the final baiss of each algorithm is also presented. Four clusters in the data are detected by SA while the polish step produces a clearer cut on the edge..}\label{fig:polish}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \code{boa5}. Optimiser PD fails to optimise this non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
\end{figure}
\end{Schunk}

@@

\caption[Comparing 20 search paths in the PCA-projected basis space with two optimisers]{Comparing 20 search paths in the PCA-projected basis space with two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7. The optimisation is done on the six-variable data, \code{boa6} with 1D projecton index $I^{nk}(n)$. A local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*). The final basis is larger and more opaque than the start basis in each path and color represents whether the global or local optimum is found, after polishing.}\label{fig:kol-result}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis. Optimisation is conducted on the 1D projection index, $I^{nk}(n)$, in boa6 data using SA with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated on the right half of the basis space and the interpolator crosses the origin to perform the interpolation, wbhich is undesirable. An orientation check is implemented to correct the sign of the target basis if the it has an opposite oreintation to the current basis. The right panel shows the same interpolation after implementing the orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
\end{figure}
\end{Schunk}
